---
title: "Hedonic Regression: Does the Property Market Price Climate Risk and Resilience?"
author: "Dave Schaaf, Chris Clever, Neeraj Athani"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 5
)
```

## Overview

Descriptive cross-county hedonic regression of Zillow Home Value Index (ZHVI) on FEMA
National Risk Index (NRI) components, controlling for ACS demographics and state +
month fixed effects. The goal is to document whether higher climate risk and lower
community resilience are associated with lower property values, conditional on
observables.

This is a **descriptive correlation**, not a causal estimate. Coastal amenity confounds
the risk signal at the county level even with the controls included here. The event
study (notebook `05`) is the complementary causal analysis.

## 1. Setup

```{r packages}
# install.packages(c("fixest", "tidycensus", "dplyr", "readr", "tidyr",
#                     "ggplot2", "lubridate", "modelsummary", "broom"))

library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(fixest)
library(tidycensus)
library(ggplot2)
library(modelsummary)
library(readxl)
```

```{r census-key, eval=FALSE}
# One-time: register at https://api.census.gov/data/key_signup.html
# Then save your key to ~/.Renviron with:
# tidycensus::census_api_key("18e688eb703b0fd7372efbcda03d554928bbba31", install = TRUE)
# Restart R afterward.
```

## 2. Load Python-prepped panel

```{r load-data}
hed <- read_csv("../data/processed/hedonic_dataset.csv",
                col_types = cols(
                  stcofips        = col_character(),
                  state           = col_character(),
                  stateabbrv      = col_character(),
                  county          = col_character(),
                  month           = col_date(),
                  month_id        = col_character(),
                  year            = col_integer(),
                  month_num       = col_integer(),
                  nri_vintage     = col_integer(),
                  cdc_svi_regime  = col_integer(),
                  zhvi            = col_double(),
                  eal_valt        = col_double(),
                  risk_value      = col_double(),
                  crf_value       = col_double(),
                  resl_value      = col_double(),
                  resl_score      = col_double(),
                  sovi_score      = col_double(),
                  sovi_value      = col_double(),
                  population      = col_double(),
                  area            = col_double(),
                  pop_density     = col_double(),
                  coastal         = col_integer()
                ))

cat("Rows:               ", format(nrow(hed), big.mark = ","), "\n")
cat("Counties:           ", n_distinct(hed$stcofips), "\n")
cat("Months:             ", n_distinct(hed$month), "\n")
cat("Window:             ", format(min(hed$month)), "to", format(max(hed$month)), "\n")
cat("Coastal counties:   ", n_distinct(hed$stcofips[hed$coastal == 1]), "\n")

# Data type check (only sovi_value should have NAs)
na_summary <- sapply(hed, function(x) sum(is.na(x)))
cat("\nNAs by column (sovi_value expected ~55% null pre-2023, others 0):\n")
print(na_summary)
```

## 3. Pull ACS controls and USDA urban/rural designation

We need county-level controls that the NRI components don't already capture:

- **Median household income** - the largest non-NRI driver of housing value
- **Coastal County Status** - account for coastal amenity and/or risk

Considered but left off due to overlap with community resilience and social vulnerability calculations:
- **Share with bachelor's degree+** - human capital, picks up labor market quality
- **Share owner-occupied** - tenure mix shifts price dynamics
- **Median age** - lifecycle/retiree composition, related to migration patterns
- **Median year structure built** - housing stock vintage, captures durability

We deliberately do *not* pull poverty, vehicle access, household size, or disability
share - those are inputs to the CDC SVI and would create multicollinearity with
`sovi_score`.

```{r acs-pull, cache=TRUE}
# Variables of interest from 2019-2023 5-year ACS
acs_vars <- c(
  median_hh_income    = "B19013_001",  # median household income
  pop_25plus          = "B15003_001",  # denominator for education
  bach_or_higher      = "B15003_022",  # bachelor's
  master_              = "B15003_023",
  prof_                = "B15003_024",
  doctorate_           = "B15003_025",
  occ_units           = "B25003_001",  # housing denominator
  owner_units         = "B25003_002",  # owner-occupied
  median_age          = "B01002_001",
  median_yr_built     = "B25035_001"
)

acs_raw <- get_acs(
  geography = "county",
  variables = acs_vars,
  year      = 2023,
  survey    = "acs5",
  output    = "wide",
  cache_table = TRUE
)

cat("ACS counties pulled:", nrow(acs_raw), "\n")
```

```{r}
# Direct download from USDA ERS
# rucc_url <- "https://www.ers.usda.gov/sites/default/files/_laserfiche/DataFiles/53251/Ruralurbancontinuumcodes2023.xlsx"
rucc_file <- "../data/raw/rucc.xlsx"
# download.file(rucc_url, destfile = rucc_file, mode = "wb")
```

```{r}
rucc <- read_excel(rucc_file, sheet = 1) %>%
  transmute(
    stcofips = sprintf("%05d", as.integer(FIPS)),
    rucc_2023 = as.integer(RUCC_2023),
    # metro_nonmetro = ifelse(rucc_2023 <= 3, "metro", "nonmetro") make binary
  )

hed <- hed %>% inner_join(rucc, by = "stcofips")
```


```{r acs-tidy}
acs <- acs_raw %>%
  transmute(
    stcofips         = GEOID,
    median_hh_income = median_hh_incomeE,
    share_bach_plus  = (bach_or_higherE + master_E + prof_E + doctorate_E) / pop_25plusE,
    share_owner_occ  = owner_unitsE / occ_unitsE,
    median_age       = median_ageE,
    median_yr_built  = median_yr_builtE
  ) %>%
  filter(!is.na(median_hh_income))

# Save a cached copy so we don't hammer the API on every render
saveRDS(acs, "../data/processed/acs_controls.rds")

summary(acs)
```

## 4. Merge and finalize variables

```{r merge}
df <- hed %>%
  inner_join(acs, by = "stcofips") %>%
  mutate(
    # Outcome
    log_zhvi          = log(zhvi),
    # NRI in logs (multiplicative form: log(risk) = log(eal) + log(crf))
    log_eal           = log(eal_valt + 1),
    log_crf           = log(crf_value),
    log_risk          = log(risk_value + 1),
    log_resl_value    = log(resl_value),
    # ACS controls in logs / standardized
    log_income        = log(median_hh_income),
    log_pop_density   = log(pmax(pop_density, 1e-3)),
    # Standardized scores for interpretable coefficients
    sovi_score_z      = as.numeric(scale(sovi_score)),
    resl_score_z      = as.numeric(scale(resl_score)),
    resl_value_z      = as.numeric(scale(resl_value))
  ) %>%
  filter(is.finite(log_zhvi), is.finite(log_eal), is.finite(log_resl_value))

cat("Final analytic rows:", format(nrow(df), big.mark = ","), "\n")
cat("Counties:           ", n_distinct(df$stcofips), "\n")
```

```{r}
df %>%
  distinct(stcofips, .keep_all = TRUE) %>%
  group_by(rucc_2023) %>%
  summarise(
    n_counties     = n(),
    mean_resl      = mean(resl_value, na.rm = TRUE),
    mean_zhvi      = mean(zhvi, na.rm = TRUE),
    mean_income    = mean(median_hh_income, na.rm = TRUE)
  )
```


## 5. Model specifications

We run three specifications side by side. All use **state + month fixed effects** with
standard errors clustered at the county level.

- **Spec A (controls):** ZHVI ~ log(income) + log(pop_density) + coastal
- **Spec B (parsimonious, NRI's intended decomposition):** ZHVI ~ log(EAL) + log(CRF)
- **Spec C (separates resilience from social vulnerability):** ZHVI ~ log(EAL) + resilience + SoVI
- **Spec D (Spec B + regime interaction):** allow SoVI coefficient to differ pre/post March 2023
- **Spec E (Spec B - Sovi):** check if resilience multicollinearity with sovi impacts coefficient
- **Spec F (Spec B - income):** check to see what income is absorbing from the NRI values
  to flag whether the HVRI→CDC methodology change shifts the apparent SoVI premium
- **Spec G (Spec B + urban-rural continuum): *** see if resilience coefficient changes when urban/ruralness accounted for

```{r specs}
# Note: removed controls included in community resilience: share_bach_plus, share_owner_occ. Age partially included in social vulnerability calculation. Income also partially included, but keeping it as a separate control
ctrls <- "log_income + log_pop_density + coastal"
fe    <- "state + month_id"

spec_A <- as.formula(paste(
  "log_zhvi ~ ", ctrls, "|", fe
))

spec_B <- as.formula(paste(
  "log_zhvi ~ log_eal + log_crf +", ctrls, "|", fe
))

spec_C <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +", ctrls, "|", fe
))

spec_D <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z + sovi_score_z:cdc_svi_regime +",
  ctrls, "|", fe
))

spec_E <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z +",
  ctrls, "|", fe
))

ctrls <- "log_pop_density + coastal"
spec_F <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +",
  ctrls, "|", fe
))

# add urban/rural continuum to fixed effects
ctrls <- "log_income + log_pop_density + coastal"
fe <- "state + month_id + factor(rucc_2023)"
spec_G <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +", 
  ctrls, "|", fe
))
```

```{r fit}
mA <- feols(spec_A, data = df, cluster = ~ stcofips)
mB <- feols(spec_B, data = df, cluster = ~ stcofips)
mC <- feols(spec_C, data = df, cluster = ~ stcofips)
mD <- feols(spec_D, data = df, cluster = ~ stcofips)
mE <- feols(spec_E, data = df, cluster = ~ stcofips)
mF <- feols(spec_F, data = df, cluster = ~ stcofips)
mG <- feols(spec_G, data = df, cluster = ~ stcofips)

etable(mA, mB, mC, mD, mE, mF, mG,
       title = "Hedonic regression: log(ZHVI) on NRI components and controls",
       digits  = 3,
       fitstat = c("n", "r2", "war2"))
```

```{r}
# Urban/rural control comparison
etable(mB, mF, headers = c("No RUCC", "RUCC FE"), digits = 3)
```


## 6. Interpretation guide

For continuous logged predictors (`log_eal`, `log_crf`, `log_income`, `log_pop_density`):
each coefficient is approximately the percent change in ZHVI per 1% change in the predictor
(elasticity).

For standardized predictors (`resl_value_z`, `sovi_score_z`): coefficients are the percent
change in ZHVI per **1 standard deviation** increase in the predictor.

For the binary `coastal` and the `share_*` and `median_age` controls: coefficients are
percent change in ZHVI per unit change.

**Expected directions if markets price climate risk and resilience:**

- `log_eal` < 0: higher expected losses → lower prices
- `log_crf` < 0: higher community risk factor → lower prices
- `resl_value_z` > 0: more resilient community → higher prices
- `sovi_score_z` likely < 0 but interpretation is mostly about the population, not climate
- `coastal` > 0 (and probably large): the amenity premium swamps the risk discount in raw form

The regime-interaction term in Spec C tests whether the SoVI coefficient shifts when
the underlying SoVI methodology changes. A large, significant interaction means we
can't interpret the pooled SoVI coefficient cleanly.

## 7. Robustness

```{r robustness}
# R1: drop coastal counties entirely (test whether non-coastal risk has its own pricing)
m_nocoast <- feols(spec_C,
                   data = df %>% filter(coastal == 0),
                   cluster = ~ stcofips)

# R2: state-clustered SEs (more conservative)
m_state_cluster <- feols(spec_C, data = df, cluster = ~ state)

# R3: 2023+ subsample only (CDC SVI regime, single methodology)
m_2023plus <- feols(spec_C,
                    data = df %>% filter(cdc_svi_regime == 1),
                    cluster = ~ stcofips)

# R4: 2020-2022 subsample only (HVRI SoVI regime)
m_pre2023 <- feols(spec_C,
                   data = df %>% filter(cdc_svi_regime == 0),
                   cluster = ~ stcofips)

etable(mC, m_nocoast, m_state_cluster, m_2023plus, m_pre2023,
       headers = c("Main", "Inland only", "State cluster", "CDC regime", "HVRI regime"),
       title   = "Robustness checks against Spec B",
       digits  = 3,
       fitstat = c("n", "r2"))
```

## 7b. Cross-section robustness: one observation per county per NRI vintage

Loads the collapsed dataset built in `06_hedonic.ipynb`: one row per county per NRI
vintage, with mean ZHVI over each vintage's coverage window. Sample drops from ~189k
to ~9k.

**Why this matters:**

- **Sidesteps the time-invariance issue with ACS controls.** In the panel, our 2019-2023
  ACS income is matched against ZHVI from 2020-2025 - real income changed over those
  years, so the static income variable carries measurement error that attenuates the
  income coefficient and lets correlated variables (resilience, coastal, density)
  absorb the missing income trend. Collapsing to one observation per vintage with mean
  ZHVI over a window inside the ACS span eliminates the mismatch.
- **Removes within-county serial correlation.** County clustering in the panel is
  designed to handle this, but a true cross-section avoids the issue entirely.
- **Matches resolution of NRI updates.** NRI is a vintage-anchored measurement, not a
  smoothly varying signal. The cross-section respects that data structure.

If point estimates are similar to the panel, that's evidence the panel findings
are not artifacts of time-aggregation choices. If they differ meaningfully, the
cross-section should be preferred.

```{r xsec-load}
xsec <- read_csv("../data/processed/hedonic_xsec.csv",
                 col_types = cols(
                   stcofips        = col_character(),
                   state           = col_character(),
                   stateabbrv      = col_character(),
                   county          = col_character(),
                   window_start    = col_date(),
                   window_end      = col_date(),
                   n_months        = col_integer(),
                   nri_vintage     = col_integer(),
                   cdc_svi_regime  = col_integer(),
                   coastal         = col_integer(),
                   .default        = col_double()
                 ))

cat("Cross-section rows: ", nrow(xsec), "\n")
cat("Counties per vintage:\n")
print(table(xsec$nri_vintage))
```

### Income data for the cross-section

**ACS 5-year, vintage-matched** is the median income source because it matches the analysis approach used in most published
hedonic real estate work. Vintage window-end aligned with NRI release (*one year lag for 2025 as 2021-2025 data not yet available).

```{r xsec-acs-vintage-matched, cache=TRUE}
# Pull ACS for each vintage and stack
fetch_acs <- function(end_year) {
  get_acs(
    geography = "county",
    variables = c(median_hh_income = "B19013_001"),
    year      = end_year,
    survey    = "acs5",
    output    = "wide",
    cache_table = TRUE
  ) %>%
    transmute(
      stcofips         = GEOID,
      median_hh_income = median_hh_incomeE,
      acs_end_year     = end_year
    ) %>%
    filter(!is.na(median_hh_income))
}

acs_xs <- bind_rows(
  fetch_acs(2020) %>% mutate(nri_vintage = 2020L),
  fetch_acs(2021) %>% mutate(nri_vintage = 2021L),
  fetch_acs(2023) %>% mutate(nri_vintage = 2023L),
  fetch_acs(2024) %>% mutate(nri_vintage = 2025L)
) %>%
  select(stcofips, nri_vintage, median_hh_income, acs_end_year)

cat("ACS vintage-matched rows:", nrow(acs_xs), "\n")
```

```{r xsec-merge}
# Counts before ACS join (already filtered upstream in 06_hedonic.ipynb for NRI∩ZHVI)
pre_acs_counts <- xsec %>%
  count(nri_vintage, name = "n_pre_acs")

xsec_joined <- xsec %>%
  inner_join(acs_xs, by = c("stcofips", "nri_vintage"))

# Counts after ACS join but before finite-log filter
post_acs_counts <- xsec_joined %>%
  count(nri_vintage, name = "n_post_acs")

xsec_df <- xsec_joined %>%
  mutate(
    log_zhvi        = log(zhvi),
    log_eal         = log(eal_valt + 1),
    log_crf         = log(crf_value),
    log_resl_value  = log(resl_value),
    log_income      = log(median_hh_income),
    log_pop_density = log(pmax(pop_density, 1e-3)),
    sovi_score_z    = as.numeric(scale(sovi_score)),
    resl_value_z    = as.numeric(scale(resl_value))
  ) %>%
  filter(is.finite(log_zhvi), is.finite(log_eal), is.finite(log_resl_value))

# Final counts after finite-log filter
post_filter_counts <- xsec_df %>%
  count(nri_vintage, name = "n_final")

# Attrition ledger: NRI∩ZHVI → +ACS → +finite-logs
attrition_ledger <- pre_acs_counts %>%
  left_join(post_acs_counts,    by = "nri_vintage") %>%
  left_join(post_filter_counts, by = "nri_vintage") %>%
  mutate(
    drop_acs    = n_pre_acs - n_post_acs,
    drop_logs   = n_post_acs - n_final
  )

cat("Cross-section attrition by vintage:\n")
print(attrition_ledger)
cat("\nFinal cross-section rows:", nrow(xsec_df), "\n")
```

The `n_pre_acs` column reflects counties present after the upstream NRI ∩ ZHVI inner
join in `06_hedonic.ipynb`. The `drop_acs` column is the additional attrition from
the vintage-matched ACS join (counties without ACS median household income for that
vintage's end-year). The `drop_logs` column is attrition from any non-finite log
transform of ZHVI, EAL, or resilience. See `methodology.md` §2.6 for known sample
limitations (Connecticut excluded entirely; ~17–18 Alaska boroughs missing per
vintage; rural-county Zillow coverage expanding over time).

```{r xsec-fit}
# State + vintage FE.
# Vintage FE absorbs national-level differences across NRI releases (incl. CDC vs HVRI SoVI regime).
# State FE absorbs everything time-invariant about each state.
fe_xs <- "state + nri_vintage"

xs_A <- feols(
  as.formula(paste(
    "log_zhvi ~ log_eal + log_crf + log_income + log_pop_density + coastal |", fe_xs
  )),
  data = xsec_df, cluster = ~ stcofips
)

xs_B <- feols(
  as.formula(paste(
    "log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal |", fe_xs
  )),
  data = xsec_df, cluster = ~ stcofips
)

# Single-vintage cuts: cleanest possible cross-section (no methodology mixing)
xs_2025 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec_df %>% filter(nri_vintage == 2025),
  cluster = ~ stcofips
)
xs_2023 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec_df %>% filter(nri_vintage == 2023),
  cluster = ~ stcofips
)
xs_2021 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec_df %>% filter(nri_vintage == 2021),
  cluster = ~ stcofips
)
xs_2020 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec_df %>% filter(nri_vintage == 2020),
  cluster = ~ stcofips
)
```

```{r xsec-compare}
etable(mC, xs_A, xs_B, xs_2020, xs_2021, xs_2023, xs_2025,
       headers = c("Panel", "XSec all basic", "XSec all", "XSec 2020", "XSec 2021", "XSec 2023", "XSec 2025"),
       title   = "Panel hedonic vs vintage-collapsed cross-section",
       digits  = 3,
       fitstat = c("n", "r2"))
```

**What to look for:**

- If `log_eal`, `resl_value_z`, and `sovi_score_z` coefficients hold sign and rough
  magnitude across panel and cross-section, the panel findings are robust.
- SEs widen because n drops from ~189k to ~9k (or ~3k for single-vintage cuts) -
  expected and not a problem unless coefficients also change sign.
- The 2023 cross-section is the cleanest single-vintage cut: CDC SVI throughout,
  ACS 2019-2023 income matched to mean ZHVI Mar 2023 - Nov 2025, no time variation
  in any predictor.

## 8. Implied-discount-rate calibration (back-of-envelope)

If a 1% increase in expected annual loss reduces ZHVI by β₁ percent, then β₁ implies the
market's effective capitalization rate on hazard losses. Compare to a 5% benchmark
(rough perpetuity discount on annual losses).

```{r implied-rate}
beta_eal <- coef(mB)["log_eal"]

# Median ZHVI and median EAL in the sample
med_zhvi <- median(df$zhvi)
med_eal  <- median(df$eal_valt)

# Implied $ change in ZHVI from a $1 change in EAL (at the medians)
implied_dollar <- beta_eal * med_zhvi / med_eal

cat("β on log(EAL):                       ", round(beta_eal, 4), "\n")
cat("Implied $-discount per $ of annual EAL:", round(implied_dollar, 2), "\n")
cat("Benchmark (5% rate, perpetuity):       -20.00\n")
cat("Example: ", -0.00225 * 300000 / 4000000,"\n")
cat("\nIf the implied figure is ~-20, the market is pricing risk near a 5% rate.\n")
cat("If it's ~ -2, the market is pricing risk near a 50% rate (heavy discount).\n")
cat("If it's near 0 or positive, the market is essentially ignoring risk.\n")
```

## 9. Caveats and what this analysis cannot establish

This regression measures *correlation conditional on observables*. The county-level
coastal dummy and ACS controls do not fully purge amenity bias (people pay more to
live near oceans regardless of climate risk). The hedonic coefficient is therefore a
**lower bound on the true risk discount in absolute terms** - the market may discount
risk more than we measure, but the amenity premium is netting against it.

The SoVI methodology break in March 2023 means the pooled-sample SoVI coefficient
mixes two different measurements. The regime-interaction term in Spec C and the
stratified robustness checks test whether this matters in practice.

ZHVI is smoothed and seasonally adjusted. For the cross-sectional question this
matters less than for the event study, but month-to-month price discovery is
attenuated.

The event study (notebook `05`) provides a complementary, less amenity-confounded
test of whether markets *update* when risk is realized via storm.
