---
title: "Hedonic Regression: Does the Property Market Price Climate Risk and Resilience?"
author: "Dave Schaaf, Chris Clever, Neeraj Athani"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 5
)
```

## Overview

Descriptive cross-county hedonic regression of Zillow Home Value Index (ZHVI) on FEMA
National Risk Index (NRI) components, controlling for ACS demographics and state +
month fixed effects. The goal is to document whether higher climate risk and lower
community resilience are associated with lower property values, conditional on
observables.

This is a **descriptive correlation**, not a causal estimate. Coastal amenity confounds
the risk signal at the county level even with the controls included here. The event
study (notebook `05`) is the complementary causal analysis.

## 1. Setup

```{r packages}
# install.packages(c("fixest", "tidycensus", "dplyr", "readr", "tidyr",
#                     "ggplot2", "lubridate", "modelsummary", "broom"))

library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(fixest)
library(tidycensus)
library(ggplot2)
library(modelsummary)
```

```{r census-key, eval=FALSE}
# One-time: register at https://api.census.gov/data/key_signup.html
# Then save your key to ~/.Renviron with:
# tidycensus::census_api_key("18e688eb703b0fd7372efbcda03d554928bbba31", install = TRUE)
# Restart R afterward.
```

## 2. Load Python-prepped panel

```{r load-data}
hed <- read_csv("../data/processed/hedonic_dataset.csv",
                col_types = cols(
                  stcofips        = col_character(),
                  state           = col_character(),
                  stateabbrv      = col_character(),
                  county          = col_character(),
                  month           = col_date(),
                  month_id        = col_character(),
                  year            = col_integer(),
                  month_num       = col_integer(),
                  nri_vintage     = col_integer(),
                  cdc_svi_regime  = col_integer(),
                  zhvi            = col_double(),
                  eal_valt        = col_double(),
                  risk_value      = col_double(),
                  crf_value       = col_double(),
                  resl_value      = col_double(),
                  resl_score      = col_double(),
                  sovi_score      = col_double(),
                  sovi_value      = col_double(),
                  population      = col_double(),
                  area            = col_double(),
                  pop_density     = col_double(),
                  coastal         = col_integer()
                ))

cat("Rows:               ", format(nrow(hed), big.mark = ","), "\n")
cat("Counties:           ", n_distinct(hed$stcofips), "\n")
cat("Months:             ", n_distinct(hed$month), "\n")
cat("Window:             ", format(min(hed$month)), "to", format(max(hed$month)), "\n")
cat("Coastal counties:   ", n_distinct(hed$stcofips[hed$coastal == 1]), "\n")

# Data type check (only sovi_value should have NAs)
na_summary <- sapply(hed, function(x) sum(is.na(x)))
cat("\nNAs by column (sovi_value expected ~55% null pre-2023, others 0):\n")
print(na_summary)
```

## 3. Pull ACS controls and USDA urban/rural designation

We need county-level controls that the NRI components don't already capture:

- **Median household income** - the largest non-NRI driver of housing value
- **Coastal County Status** - account for coastal amenity and/or risk

Considered but left off due to overlap with community resilience and social vulnerability calculations:
- **Share with bachelor's degree+** - human capital, picks up labor market quality
- **Share owner-occupied** - tenure mix shifts price dynamics
- **Median age** - lifecycle/retiree composition, related to migration patterns
- **Median year structure built** - housing stock vintage, captures durability

We deliberately do *not* pull poverty, vehicle access, household size, or disability
share - those are inputs to the CDC SVI and would create multicollinearity with
`sovi_score`.

```{r acs-pull, cache=TRUE}
# Variables of interest from 2019-2023 5-year ACS
acs_vars <- c(
  median_hh_income    = "B19013_001",  # median household income
  pop_25plus          = "B15003_001",  # denominator for education
  bach_or_higher      = "B15003_022",  # bachelor's
  master_              = "B15003_023",
  prof_                = "B15003_024",
  doctorate_           = "B15003_025",
  occ_units           = "B25003_001",  # housing denominator
  owner_units         = "B25003_002",  # owner-occupied
  median_age          = "B01002_001",
  median_yr_built     = "B25035_001"
)

acs_raw <- get_acs(
  geography = "county",
  variables = acs_vars,
  year      = 2023,
  survey    = "acs5",
  output    = "wide",
  cache_table = TRUE
)

cat("ACS counties pulled:", nrow(acs_raw), "\n")
```

```{r}
# Direct download from USDA ERS
rucc_url <- "https://www.ers.usda.gov/sites/default/files/_laserfiche/DataFiles/53251/Ruralurbancontinuumcodes2023.xlsx"
rucc_file <- "../data/raw/rucc.xlsx"
download.file(rucc_url, destfile = rucc_file, mode = "wb")
```

```{r}
library(readxl)
rucc <- read_excel(rucc_file, sheet = 1) %>%
  transmute(
    stcofips = sprintf("%05d", as.integer(FIPS)),
    rucc_2023 = as.integer(RUCC_2023),
    # metro_nonmetro = ifelse(rucc_2023 <= 3, "metro", "nonmetro") make binary
  )

hed <- hed %>% inner_join(rucc, by = "stcofips")
```


```{r acs-tidy}
acs <- acs_raw %>%
  transmute(
    stcofips         = GEOID,
    median_hh_income = median_hh_incomeE,
    share_bach_plus  = (bach_or_higherE + master_E + prof_E + doctorate_E) / pop_25plusE,
    share_owner_occ  = owner_unitsE / occ_unitsE,
    median_age       = median_ageE,
    median_yr_built  = median_yr_builtE
  ) %>%
  filter(!is.na(median_hh_income))

# Save a cached copy so we don't hammer the API on every render
saveRDS(acs, "../data/processed/acs_controls.rds")

summary(acs)
```

## 4. Merge and finalize variables

```{r merge}
df <- hed %>%
  inner_join(acs, by = "stcofips") %>%
  mutate(
    # Outcome
    log_zhvi          = log(zhvi),
    # NRI in logs (multiplicative form: log(risk) = log(eal) + log(crf))
    log_eal           = log(eal_valt + 1),
    log_crf           = log(crf_value),
    log_risk          = log(risk_value + 1),
    log_resl_value    = log(resl_value),
    # ACS controls in logs / standardized
    log_income        = log(median_hh_income),
    log_pop_density   = log(pmax(pop_density, 1e-3)),
    # Standardized scores for interpretable coefficients
    sovi_score_z      = as.numeric(scale(sovi_score)),
    resl_score_z      = as.numeric(scale(resl_score)),
    resl_value_z      = as.numeric(scale(resl_value))
  ) %>%
  filter(is.finite(log_zhvi), is.finite(log_eal), is.finite(log_resl_value))

cat("Final analytic rows:", format(nrow(df), big.mark = ","), "\n")
cat("Counties:           ", n_distinct(df$stcofips), "\n")
```

```{r}
df %>%
  distinct(stcofips, .keep_all = TRUE) %>%
  group_by(rucc_2023) %>%
  summarise(
    n_counties     = n(),
    mean_resl      = mean(resl_value, na.rm = TRUE),
    mean_zhvi      = mean(zhvi, na.rm = TRUE),
    mean_income    = mean(median_hh_income, na.rm = TRUE)
  )
```


## 5. Model specifications

We run three specifications side by side. All use **state + month fixed effects** with
standard errors clustered at the county level.

- **Spec A.1 (controls):** ZHVI ~ income + pop + coastal
- **Spec A.2 (parsimonious, NRI's intended decomposition):** ZHVI ~ log(EAL) + log(CRF)
- **Spec B (separates resilience from social vulnerability):** ZHVI ~ log(EAL) + resilience + SoVI
- **Spec C (Spec B + regime interaction):** allow SoVI coefficient to differ pre/post March 2023
- **Spec D (Spec B - Sovi):** check if resilience multicollinearity with sovi impacts coefficient
- **Spec E (Spec B - income):** check to see what income is absorbing from the NRI values
  to flag whether the HVRI→CDC methodology change shifts the apparent SoVI premium
- **Spec G (Spec B + urban-rural continuum): *** see if resilience coefficient changes when urban/ruralness accounted for

```{r specs}
# Note: removed controls included in community resilience: share_bach_plus, share_owner_occ. Age partially included in social vulnerability calculation. Income also partially included, but keeping it as a separate control
ctrls <- "log_income + log_pop_density + coastal"
fe    <- "state + month_id"

spec_A1 <- as.formula(paste(
  "log_zhvi ~ ", ctrls, "|", fe
))

spec_A2 <- as.formula(paste(
  "log_zhvi ~ log_eal + log_crf +", ctrls, "|", fe
))

spec_B <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +", ctrls, "|", fe
))

spec_C <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z + sovi_score_z:cdc_svi_regime +",
  ctrls, "|", fe
))

spec_D <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z +",
  ctrls, "|", fe
))

ctrls <- "log_pop_density + coastal"
spec_E <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +",
  ctrls, "|", fe
))

# add urban/rural continuum to fixed effects
ctrls <- "log_income + log_pop_density + coastal"
fe <- "state + month_id + factor(rucc_2023)"
spec_F <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +", 
  ctrls, "|", fe
))
```

```{r fit}
mA1 <- feols(spec_A1, data = df, cluster = ~ stcofips)
mA2 <- feols(spec_A2, data = df, cluster = ~ stcofips)
mB <- feols(spec_B, data = df, cluster = ~ stcofips)
mC <- feols(spec_C, data = df, cluster = ~ stcofips)
mD <- feols(spec_D, data = df, cluster = ~ stcofips)
mE <- feols(spec_E, data = df, cluster = ~ stcofips)
mF <- feols(spec_F, data = df, cluster = ~ stcofips)

etable(mA1, mA2, mB, mC, mD, mE, mF, 
       title = "Hedonic regression: log(ZHVI) on NRI components and controls",
       digits  = 3,
       fitstat = c("n", "r2", "war2"))
```

```{r}
# Urban/rural control comparison
etable(mB, mF, headers = c("No RUCC", "RUCC FE"), digits = 3)
```


## 6. Interpretation guide

For continuous logged predictors (`log_eal`, `log_crf`, `log_income`, `log_pop_density`):
each coefficient is approximately the percent change in ZHVI per 1% change in the predictor
(elasticity).

For standardized predictors (`resl_value_z`, `sovi_score_z`): coefficients are the percent
change in ZHVI per **1 standard deviation** increase in the predictor.

For the binary `coastal` and the `share_*` and `median_age` controls: coefficients are
percent change in ZHVI per unit change.

**Expected directions if markets price climate risk and resilience:**

- `log_eal` < 0: higher expected losses → lower prices
- `log_crf` < 0: higher community risk factor → lower prices
- `resl_value_z` > 0: more resilient community → higher prices
- `sovi_score_z` likely < 0 but interpretation is mostly about the population, not climate
- `coastal` > 0 (and probably large): the amenity premium swamps the risk discount in raw form

The regime-interaction term in Spec C tests whether the SoVI coefficient shifts when
the underlying SoVI methodology changes. A large, significant interaction means we
can't interpret the pooled SoVI coefficient cleanly.

## 7. Robustness

```{r robustness}
# R1: drop coastal counties entirely (test whether non-coastal risk has its own pricing)
m_nocoast <- feols(spec_B,
                   data = df %>% filter(coastal == 0),
                   cluster = ~ stcofips)

# R2: state-clustered SEs (more conservative)
m_state_cluster <- feols(spec_B, data = df, cluster = ~ state)

# R3: 2023+ subsample only (CDC SVI regime, single methodology)
m_2023plus <- feols(spec_B,
                    data = df %>% filter(cdc_svi_regime == 1),
                    cluster = ~ stcofips)

# R4: 2020-2022 subsample only (HVRI SoVI regime)
m_pre2023 <- feols(spec_B,
                   data = df %>% filter(cdc_svi_regime == 0),
                   cluster = ~ stcofips)

etable(mB, m_nocoast, m_state_cluster, m_2023plus, m_pre2023,
       headers = c("Main", "Inland only", "State cluster", "CDC regime", "HVRI regime"),
       title   = "Robustness checks against Spec B",
       digits  = 3,
       fitstat = c("n", "r2"))
```

## 8. Implied-discount-rate calibration (back-of-envelope)

If a 1% increase in expected annual loss reduces ZHVI by β₁ percent, then β₁ implies the
market's effective capitalization rate on hazard losses. Compare to a 5% benchmark
(rough perpetuity discount on annual losses).

```{r implied-rate}
beta_eal <- coef(mB)["log_eal"]

# Median ZHVI and median EAL in the sample
med_zhvi <- median(df$zhvi)
med_eal  <- median(df$eal_valt)

# Implied $ change in ZHVI from a $1 change in EAL (at the medians)
implied_dollar <- beta_eal * med_zhvi / med_eal

cat("β on log(EAL):                       ", round(beta_eal, 4), "\n")
cat("Implied $-discount per $ of annual EAL:", round(implied_dollar, 2), "\n")
cat("Benchmark (5% rate, perpetuity):       -20.00\n")
cat("Example: ", -0.00225 * 300000 / 4000000,"\n")
cat("\nIf the implied figure is ~-20, the market is pricing risk near a 5% rate.\n")
cat("If it's ~ -2, the market is pricing risk near a 50% rate (heavy discount).\n")
cat("If it's near 0 or positive, the market is essentially ignoring risk.\n")
```

## 9. Caveats and what this analysis cannot establish

This regression measures *correlation conditional on observables*. The county-level
coastal dummy and ACS controls do not fully purge amenity bias (people pay more to
live near oceans regardless of climate risk). The hedonic coefficient is therefore a
**lower bound on the true risk discount in absolute terms** - the market may discount
risk more than we measure, but the amenity premium is netting against it.

The SoVI methodology break in March 2023 means the pooled-sample SoVI coefficient
mixes two different measurements. The regime-interaction term in Spec C and the
stratified robustness checks test whether this matters in practice.

ZHVI is smoothed and seasonally adjusted. For the cross-sectional question this
matters less than for the event study, but month-to-month price discovery is
attenuated.

The event study (notebook `05`) provides a complementary, less amenity-confounded
test of whether markets *update* when risk is realized via storm.
