---
title: "Hedonic Regression: Does the Property Market Price Climate Risk and Resilience?"
author: "Dave Schaaf, Chris Clever, Neeraj Athani"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 5
)

knitr::opts_knit$set(root.dir = "/Users/davidschaaf/Projects/climate-exposure")
```

## Overview

Descriptive cross-county hedonic regression of Zillow Home Value Index (ZHVI) on FEMA
National Risk Index (NRI) components, controlling for ACS demographics and state +
month fixed effects. The goal is to document whether higher climate risk and lower
community resilience are associated with lower property values, conditional on
observables.

This is a **descriptive correlation**, not a causal estimate. Coastal amenity confounds
the risk signal at the county level even with the controls included here. The event
study (notebook `05`) is the complementary causal analysis.

## 1. Setup

```{r packages}
# install.packages(c("fixest", "tidycensus", "dplyr", "readr", "tidyr",
#                     "ggplot2", "lubridate", "modelsummary", "broom"))

library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(fixest)
library(tidycensus)
library(ggplot2)
library(modelsummary)
```

```{r census-key, eval=FALSE}
# One-time: register at https://api.census.gov/data/key_signup.html
# Then save your key to ~/.Renviron with:
# tidycensus::census_api_key("18e688eb703b0fd7372efbcda03d554928bbba31", install = TRUE)
# Restart R afterward.
```

## 2. Load Python-prepped panel

```{r load-data}
hed <- read_csv("data/processed/hedonic_dataset.csv",
                col_types = cols(
                  stcofips        = col_character(),
                  state           = col_character(),
                  stateabbrv      = col_character(),
                  county          = col_character(),
                  month           = col_date(),
                  month_id        = col_character(),
                  year            = col_integer(),
                  month_num       = col_integer(),
                  nri_vintage     = col_integer(),
                  cdc_svi_regime  = col_integer(),
                  zhvi            = col_double(),
                  eal_valt        = col_double(),
                  risk_value      = col_double(),
                  crf_value       = col_double(),
                  resl_value      = col_double(),
                  resl_score      = col_double(),
                  sovi_score      = col_double(),
                  sovi_value      = col_double(),
                  population      = col_double(),
                  area            = col_double(),
                  pop_density     = col_double(),
                  coastal         = col_integer()
                ))

cat("Rows:               ", format(nrow(hed), big.mark = ","), "\n")
cat("Counties:           ", n_distinct(hed$stcofips), "\n")
cat("Months:             ", n_distinct(hed$month), "\n")
cat("Window:             ", format(min(hed$month)), "to", format(max(hed$month)), "\n")
cat("Coastal counties:   ", n_distinct(hed$stcofips[hed$coastal == 1]), "\n")

# Data type check (only sovi_value should have NAs)
na_summary <- sapply(hed, function(x) sum(is.na(x)))
cat("\nNAs by column (sovi_value expected ~55% null pre-2023, others 0):\n")
print(na_summary)
```

## 3. Pull ACS controls and USDA urban/rural designation

We need county-level controls that the NRI components don't already capture:

- **Median household income** - the largest non-NRI driver of housing value
- **Coastal County Status** - account for coastal amenity and/or risk

Considered but left off due to overlap with community resilience and social vulnerability calculations:
- **Share with bachelor's degree+** - human capital, picks up labor market quality
- **Share owner-occupied** - tenure mix shifts price dynamics
- **Median age** - lifecycle/retiree composition, related to migration patterns
- **Median year structure built** - housing stock vintage, captures durability

We deliberately do *not* pull poverty, vehicle access, household size, or disability
share - those are inputs to the CDC SVI and would create multicollinearity with
`sovi_score`.

```{r acs-pull, cache=TRUE}
# Variables of interest from 2019-2023 5-year ACS
acs_vars <- c(
  median_hh_income    = "B19013_001",  # median household income
  pop_25plus          = "B15003_001",  # denominator for education
  bach_or_higher      = "B15003_022",  # bachelor's
  master_              = "B15003_023",
  prof_                = "B15003_024",
  doctorate_           = "B15003_025",
  occ_units           = "B25003_001",  # housing denominator
  owner_units         = "B25003_002",  # owner-occupied
  median_age          = "B01002_001",
  median_yr_built     = "B25035_001"
)

acs_raw <- get_acs(
  geography = "county",
  variables = acs_vars,
  year      = 2023,
  survey    = "acs5",
  output    = "wide",
  cache_table = TRUE
)

cat("ACS counties pulled:", nrow(acs_raw), "\n")
```

```{r}
# Direct download from USDA ERS
rucc_url <- "https://www.ers.usda.gov/sites/default/files/_laserfiche/DataFiles/53251/Ruralurbancontinuumcodes2023.xlsx"
rucc_file <- "data/raw/rucc.xlsx"
download.file(rucc_url, destfile = rucc_file, mode = "wb")
```

```{r}
library(readxl)
rucc <- read_excel(rucc_file, sheet = 1) %>%
  transmute(
    stcofips = sprintf("%05d", as.integer(FIPS)),
    rucc_2023 = as.integer(RUCC_2023),
    # metro_nonmetro = ifelse(rucc_2023 <= 3, "metro", "nonmetro") make binary
  )
rucc_2023 <- rucc['rucc_2023']
hed <- hed %>% inner_join(rucc, by = "stcofips")


```


```{r acs-tidy}
acs <- acs_raw %>%
  transmute(
    stcofips         = GEOID,
    median_hh_income = median_hh_incomeE,
    share_bach_plus  = (bach_or_higherE + master_E + prof_E + doctorate_E) / pop_25plusE,
    share_owner_occ  = owner_unitsE / occ_unitsE,
    median_age       = median_ageE,
    median_yr_built  = median_yr_builtE
  ) %>%
  filter(!is.na(median_hh_income))

# Save a cached copy so we don't hammer the API on every render
#saveRDS(acs, "../data/processed/acs_controls.rds")

summary(acs)
```

## 4. Merge and finalize variables

```{r merge}
df <- hed %>%
  inner_join(acs, by = "stcofips") %>%
  mutate(
    # Outcome
    log_zhvi          = log(zhvi),
    # NRI in logs (multiplicative form: log(risk) = log(eal) + log(crf))
    log_eal           = log(eal_valt + 1),
    log_crf           = log(crf_value),
    log_risk          = log(risk_value + 1),
    log_resl_value    = log(resl_value),
    # ACS controls in logs / standardized
    log_income        = log(median_hh_income),
    log_pop_density   = log(pmax(pop_density, 1e-3)),
    # Standardized scores for interpretable coefficients
    sovi_score_z      = as.numeric(scale(sovi_score)),
    resl_score_z      = as.numeric(scale(resl_score)),
    resl_value_z      = as.numeric(scale(resl_value))
  ) %>%
  filter(is.finite(log_zhvi), is.finite(log_eal), is.finite(log_resl_value))

cat("Final analytic rows:", format(nrow(df), big.mark = ","), "\n")
cat("Counties:           ", n_distinct(df$stcofips), "\n")
```

```{r}
df %>%
  distinct(stcofips, .keep_all = TRUE) %>%
  group_by(rucc_2023) %>%
  summarise(
    n_counties     = n(),
    mean_resl      = mean(resl_value, na.rm = TRUE),
    mean_zhvi      = mean(zhvi, na.rm = TRUE),
    mean_income    = mean(median_hh_income, na.rm = TRUE)
  )
```


## 5. Model specifications

We run six specifications side by side. All use **state + month fixed effects** with
standard errors clustered at the county level.

- **Spec A (parsimonious, NRI's intended decomposition):** ZHVI ~ log(EAL) + log(CRF)
- **Spec B (separates resilience from social vulnerability):** ZHVI ~ log(EAL) + resilience + SoVI
- **Spec C (Spec B + regime interaction):** allow SoVI coefficient to differ pre/post March 2023
- **Spec D (Spec B - Sovi):** check if resilience multicollinearity with sovi impacts coefficient
- **Spec E (Spec B - income):** check to see what income is absorbing from the NRI values
  to flag whether the HVRI→CDC methodology change shifts the apparent SoVI premium
- **Spec F (Spec B + urban-rural continuum): *** see if resilience coefficient changes when urban/ruralness accounted for

```{r specs}
# Note: removed controls included in community resilience: share_bach_plus, share_owner_occ. Age partially included in social vulnerability calculation. Income also partially included, but keeping it as a separate control
ctrls <- "log_income + log_pop_density + coastal"
fe    <- "state + month_id"

spec_A <- as.formula(paste(
  "log_zhvi ~ log_eal + log_crf +", ctrls, "|", fe
))

spec_B <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +", ctrls, "|", fe
))

spec_C <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z + sovi_score_z:cdc_svi_regime +",
  ctrls, "|", fe
))

spec_D <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z +",
  ctrls, "|", fe
))

ctrls <- "log_pop_density + coastal"
spec_E <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +",
  ctrls, "|", fe
))

# add urban/rural continuum to fixed effects
ctrls <- "log_income + log_pop_density + coastal"
fe <- "state + month_id + factor(rucc_2023)"
spec_F <- as.formula(paste(
  "log_zhvi ~ log_eal + resl_value_z + sovi_score_z +", 
  ctrls, "|", fe
))
```

```{r fit}
mA <- feols(spec_A, data = df, cluster = ~ stcofips)
mB <- feols(spec_B, data = df, cluster = ~ stcofips)
mC <- feols(spec_C, data = df, cluster = ~ stcofips)
mD <- feols(spec_D, data = df, cluster = ~ stcofips)
mE <- feols(spec_E, data = df, cluster = ~ stcofips)
mF <- feols(spec_F, data = df, cluster = ~ stcofips)

  etable(mA, mB, mC, mD, mE, mF,
         title = "Hedonic regression: log(ZHVI) on NRI components and controls",
         headers = c("NRI Basic", "NRI Split", "NRI Regime", "No SOVI", "No Income", "Urban-Rural FE"),
         dict = c(
           log_zhvi                      = "Log home value (ZHVI)",
           log_income                    = "Log median income",
           log_pop_density               = "Log pop.\\ density",
           coastal                       = "Coastal county",
           log_eal                       = "Log expected annual loss",
           log_crf                       = "Log community risk factor",
           resl_value_z                  = "Resilience",
           sovi_score_z                  = "Social vulnerability",
           "sovi_score_z:cdc_svi_regime" = "SoVI $\\times$ CDC regime",
           state                         = "State",
           month_id                      = "Month",
           "factor(rucc_2023)"           = "Urban-rural (RUCC)"
         ),
         digits  = 3,
         fitstat = c("n", "r2", "war2"),
         tex=TRUE,
         file="paper/output/regression/hedonic_regression.tex")
```



```{r tbl-hedonic}
#| results: 'asis'
#| tbl-cap: "Hedonic regression specifications A, B, C, D, E"
writeLines(readLines("paper/output/regression/hedonic_regression.tex"))
```

```{r}
# Urban/rural control comparison
etable(mB, mF, headers = c("No RUCC", "RUCC FE"), digits = 3)
```


## 6. Interpretation guide

For continuous logged predictors (`log_eal`, `log_crf`, `log_income`, `log_pop_density`):
each coefficient is approximately the percent change in ZHVI per 1% change in the predictor
(elasticity).

For standardized predictors (`resl_value_z`, `sovi_score_z`): coefficients are the percent
change in ZHVI per **1 standard deviation** increase in the predictor.

For the binary `coastal` control: coefficients are
percent change in ZHVI per unit change.

**Expected directions if markets price climate risk and resilience:**

- `log_eal` < 0: higher expected losses → lower prices
- `log_crf` < 0: higher community risk factor → lower prices
- `resl_value_z` > 0: more resilient community → higher prices
- `sovi_score_z` likely < 0 but interpretation is mostly about the population, not climate
- `coastal` > 0 (and probably large): the amenity premium swamps the risk discount in raw form

The regime-interaction term in Spec C tests whether the SoVI coefficient shifts when
the underlying SoVI methodology changes. A large, significant interaction means we
can't interpret the pooled SoVI coefficient cleanly.

## 7. Robustness

```{r robustness}
# R1: drop coastal counties entirely (test whether non-coastal risk has its own pricing)
m_nocoast <- feols(spec_B,
                   data = df %>% filter(coastal == 0),
                   cluster = ~ stcofips)

# R2: state-clustered SEs (more conservative)
m_state_cluster <- feols(spec_B, data = df, cluster = ~ state)

# R3: 2023+ subsample only (CDC SVI regime, single methodology)
m_2023plus <- feols(spec_B,
                    data = df %>% filter(cdc_svi_regime == 1),
                    cluster = ~ stcofips)

# R4: 2020-2022 subsample only (HVRI SoVI regime)
m_pre2023 <- feols(spec_B,
                   data = df %>% filter(cdc_svi_regime == 0),
                   cluster = ~ stcofips)

etable(mB, m_nocoast, m_state_cluster, m_2023plus, m_pre2023,
       headers = c("Main", "Inland only", "State cluster", "CDC regime", "HVRI regime"),
       title   = "Robustness checks against Spec B",
       dict = c(
         log_zhvi        = "Log home value (ZHVI)",
         log_income      = "Log median income",
         log_pop_density = "Log pop.\\ density",
         coastal         = "Coastal county",
         log_eal         = "Log expected annual loss",
         resl_value_z    = "Resilience",
         sovi_score_z    = "Social vulnerability"
       ),
       digits  = 3,
       fitstat = c("n", "r2"),
       tex     = TRUE,
       file    = "paper/output/regression/table_robustness.tex")
```

## 7b. Cross-section robustness: one observation per county per NRI vintage

Loads the collapsed dataset built in `06_hedonic.ipynb`: one row per county per NRI
vintage, with mean ZHVI over each vintage's coverage window. Sample drops from ~189k
to ~9k.

**Purpose:**

- **Sidesteps the time-invariance issue with ACS controls.** In the panel, our 2019-2023
  ACS income is matched against ZHVI from 2020-2025 - real income changed over those
  years, so the static income variable carries measurement error that attenuates the
  income coefficient and lets correlated variables (resilience, coastal, density)
  absorb the missing income trend. Collapsing to one observation per vintage with mean
  ZHVI over a window inside the ACS span eliminates the mismatch.
- **Removes within-county serial correlation.** County clustering in the panel is
  designed to handle this, but a true cross-section avoids the issue entirely.
- **Matches resolution of NRI updates.** NRI is a vintage-anchored measurement, not a
  smoothly varying signal. The cross-section respects that data structure.

If point estimates are similar to the panel, that's evidence the panel findings
are not artifacts of time-aggregation choices. If they differ meaningfully, the
cross-section should be preferred.

```{r xsec-load}
xsec_1 <- read_csv("data/processed/hedonic_xsec.csv",
                 col_types = cols(
                   stcofips        = col_character(),
                   state           = col_character(),
                   stateabbrv      = col_character(),
                   county          = col_character(),
                   window_start    = col_date(),
                   window_end      = col_date(),
                   n_months        = col_integer(),
                   nri_vintage     = col_integer(),
                   cdc_svi_regime  = col_integer(),
                   coastal         = col_integer(),
                   .default        = col_double()
                 ))

cat("Cross-section rows: ", nrow(xsec_1), "\n")
cat("Counties per vintage:\n")
print(table(xsec_1$nri_vintage))
```

### Income data choices for the cross-section

The collapsed cross-section opens up better income data than was available in the panel.
For a cross-section, we want **county-level income matched to the vintage's coverage
window**. 

**ACS 5-year estimates from `tidycensus`, vintage-matched.**

For each NRI vintage's window, the ACS 5-year estimate that *brackets* that window:

| NRI vintage | Window | ACS 5-year vintage to use |
|---|---|---|
| 2020 | Nov 2020 - Oct 2021 | ACS 2017-2021 (centered ~2019) |
| 2021 | Nov 2021 - Feb 2023 | ACS 2018-2022 (centered ~2020) |
| 2023 | Mar 2023 - Nov 2025 | ACS 2019-2023 (centered ~2021) |
| 2025 | Nov 2025 - Mar 2026 | ACS 2020-2024 (2021-2025 not released yet) |

This is internally consistent (each ZHVI mean is matched to an income measure from
within the same window) and uses the most reliable Census product. ACS 1-year is more
current but has unusable margins of error for small counties. We don't use ACS 1-year.


```{r xsec-acs-vintage-matched, cache=TRUE}
# Pull ACS for each vintage and stack
fetch_acs_1 <- function(end_year) {
  get_acs(
    geography = "county",
    variables = c(median_hh_income = "B19013_001"),
    year      = end_year,
    survey    = "acs5",
    output    = "wide",
    cache_table = TRUE
  ) %>%
    transmute(
      stcofips         = GEOID,
      median_hh_income = median_hh_incomeE,
      acs_end_year     = end_year
    ) %>%
    filter(!is.na(median_hh_income))
}

acs_xs_1 <- bind_rows(
  fetch_acs_1(2021) %>% mutate(nri_vintage = 2020L),
  fetch_acs_1(2022) %>% mutate(nri_vintage = 2021L),
  fetch_acs_1(2023) %>% mutate(nri_vintage = 2023L),
  fetch_acs_1(2024) %>% mutate(nri_vintage = 2025L)
) %>%
  select(stcofips, nri_vintage, median_hh_income, acs_end_year)

cat("ACS vintage-matched rows:", nrow(acs_xs_1), "\n")
```

```{r xsec-merge}
xsec1_df <- xsec_1 %>%
  inner_join(acs_xs_1, by = c("stcofips", "nri_vintage")) %>%
  mutate(
    log_zhvi        = log(zhvi),
    log_eal         = log(eal_valt + 1),
    log_crf         = log(crf_value),
    log_resl_value  = log(resl_value),
    log_income      = log(median_hh_income),
    log_pop_density = log(pmax(pop_density, 1e-3)),
    sovi_score_z    = as.numeric(scale(sovi_score)),
    resl_value_z    = as.numeric(scale(resl_value))
  ) %>%
  filter(is.finite(log_zhvi), is.finite(log_eal), is.finite(log_resl_value))

cat("Final cross-section rows:", nrow(xsec1_df), "\n")
cat("By vintage:\n")
print(table(xsec1_df$nri_vintage))
```

```{r xsec-fit}
# State + vintage FE.
# Vintage FE absorbs national-level differences across NRI releases (incl. CDC vs HVRI SoVI regime).
# State FE absorbs everything time-invariant about each state.
fe_xs <- "state + nri_vintage"

xs_A <- feols(
  as.formula(paste(
    "log_zhvi ~ log_eal + log_crf + log_income + log_pop_density + coastal |", fe_xs
  )),
  data = xsec1_df, cluster = ~ stcofips
)

xs_B <- feols(
  as.formula(paste(
    "log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal |", fe_xs
  )),
  data = xsec1_df, cluster = ~ stcofips
)

# Single-vintage cuts: cleanest possible cross-section (no methodology mixing)
xs_2025 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec1_df %>% filter(nri_vintage == 2025),
  cluster = ~ stcofips
)
xs_2023 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec1_df %>% filter(nri_vintage == 2023),
  cluster = ~ stcofips
)
xs_2021 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec1_df %>% filter(nri_vintage == 2021),
  cluster = ~ stcofips
)
xs_2020 <- feols(
  log_zhvi ~ log_eal + resl_value_z + sovi_score_z + log_income + log_pop_density + coastal | state,
  data = xsec1_df %>% filter(nri_vintage == 2020),
  cluster = ~ stcofips
)
```

```{r xsec-compare}
etable(mB, xs_B, xs_2020, xs_2021, xs_2023, xs_2025,
       headers = c("Panel", "XSec all", "XSec 2020", "XSec 2021", "XSec 2023", "XSec 2025"),
       title   = "Panel hedonic vs vintage-collapsed cross-section, full ZHVI window",
       digits  = 3,
       fitstat = c("n", "r2"))
       # tex     = TRUE,
       # file    = "paper/output/regression/table_hedonic_zhvi_full_window_check.tex")

```

**What to look for:**

- If `log_eal`, `resl_value_z`, and `sovi_score_z` coefficients hold sign and rough
  magnitude across panel and cross-section, the panel findings are robust.
- SEs widen because n drops from ~189k to ~9k (or ~3k for single-vintage cuts) -
  expected and not a problem unless coefficients also change sign.
- The 2023 cross-section is the cleanest single-vintage cut: CDC SVI throughout,
  ACS 2019-2023 income matched to mean ZHVI Mar 2023 - Nov 2025, no time variation
  in any predictor.

## 8. Minimum Detectable Effect (MDE)

The MDE is the smallest true effect this design could reliably detect, given the standard errors we observed. With α = 0.05 (two-sided) and 80% power, MDE ≈ 2.8 × SE — anything smaller than that, even if real, would more often than not fail to reach significance in a study of this size.

```{r mde, message=FALSE}
library(fixest)

# Pull SEs from the pooled cross-section (xs_B). Swap in xs_2020/xs_2021/xs_2023/xs_2025
# to see how MDE changes when N drops to a single vintage.
mod <- xs_B
se_vec <- se(mod)

mde_tbl <- data.frame(
  term   = names(se_vec),
  beta   = coef(mod)[names(se_vec)],
  se     = se_vec,
  mde_80 = 2.8 * se_vec,           # 80% power, two-sided α = 0.05
  mde_90 = 3.24 * se_vec           # 90% power, two-sided α = 0.05
)
mde_tbl <- mde_tbl[mde_tbl$term %in% c("log_eal", "resl_value_z", "sovi_score_z"), ]
print(mde_tbl, row.names = FALSE)
```

**Translation to dollars.** For `log_eal`, an elasticity β implies a marginal capitalization of roughly β × (median ZHVI / median EAL) dollars of price per $1 of expected annual loss.

```{r mde-dollars}
med_zhvi <- median(xsec1_df$zhvi,    na.rm = TRUE)
med_eal  <- median(xsec1_df$eal_valt, na.rm = TRUE)
scale_factor <- med_zhvi / med_eal

mde_eal <- 2.8 * se(mod)["log_eal"]
cat(sprintf(
  "Median ZHVI: $%s; Median EAL: $%s\nMDE on log_eal (80%% power): %.4f\nImplied $-per-$1-EAL MDE: $%.5f\n5%% capitalization benchmark: $20.00 per $1 EAL\n",
  format(round(med_zhvi), big.mark = ","),
  format(round(med_eal),  big.mark = ","),
  mde_eal, mde_eal * scale_factor
))
```

**How to read it.** If the MDE on `log_eal` translates to (say) $3 per $1 of EAL, this design can rule out full 5% capitalization ($20) — the null result is informative — but it cannot distinguish "no pricing" from "small partial pricing" below the MDE threshold.

```{r hedonic-plots-rev, message=FALSE}
source(here::here("paper", "R", "06_hedonic_plots_rev.R"))
hed_plots <- run_hedonic_plots(
  df       = df,        # panel analytic frame
  xsec_df  = xsec1_df,  # vintage-collapsed cross-section
  models   = list(
    mC      = mC,                       # panel main spec (Spec C)
    mE      = mE,                       # panel, income removed
    xs_B    = xs_B,                     # cross-section main (analog of Spec C)
    xs_2020 = xs_2020, xs_2021 = xs_2021,
    xs_2023 = xs_2023, xs_2025 = xs_2025
  ),
  output_dir = here::here("paper", "output", "hedonic")
)
```

## 9. Caveats and what this analysis cannot establish

This regression measures *correlation conditional on observables*. The county-level
coastal dummy and ACS controls do not fully purge amenity bias (people pay more to
live near oceans regardless of climate risk). The hedonic coefficient is therefore a
**lower bound on the true risk discount in absolute terms** - the market may discount
risk more than we measure, but the amenity premium is netting against it.

The SoVI methodology break in March 2023 means the pooled-sample SoVI coefficient
mixes two different measurements. The regime-interaction term in Spec C and the
stratified robustness checks test whether this matters in practice.

ZHVI is smoothed and seasonally adjusted. For the cross-sectional question this
matters less than for the event study, but month-to-month price discovery is
attenuated.

The event study (notebook `05`) provides a complementary, less amenity-confounded
test of whether markets *update* when risk is realized via storm.


```{r save-hedonic-inline}
# Inline values consumed by paper/report.qmd (hedonic regression section).
# Spec labels follow THIS file:
#   A = EAL + CRF; B = resilience + SoVI; C = B + SoVI x CDC-regime interaction; D = B - SoVI.
hedonic_inline <- list(
  # Spec A — income elasticity (log-log): % change in ZHVI per 1% change in income
  hed_income_coef = round(coef(mA)["log_income"], 3),

  # Spec B — pooled social-vulnerability coefficient (per 1 SD)
  hed_sovi_coef_B = round(coef(mB)["sovi_score_z"], 3),

  # Spec C — base SoVI coefficient once the CDC-regime interaction is added
  hed_sovi_coef_C = round(coef(mC)["sovi_score_z"], 3),

  # Spec E — income removed: fragility of the remaining coefficients
  hed_eal_coef_E  = round(coef(mE)["log_eal"], 3),
  hed_resl_coef_E = round(coef(mE)["resl_value_z"], 3),
  hed_sovi_coef_E = round(coef(mE)["sovi_score_z"], 3),

  # Primary-spec log(EAL) range (A, B, C, D, F; excludes Spec E, which drops income and flips positive)
  hed_eal_min = round(min(sapply(list(mA, mB, mC, mD, mF), function(m) coef(m)["log_eal"])), 3),
  hed_eal_max = round(max(sapply(list(mA, mB, mC, mD, mF), function(m) coef(m)["log_eal"])), 3),
  mA = mA,
  mB = mB,
  mC = mC,
  mD = mD,
  mE = mE,
  mF = mF,
  rucc_df = df,
  hed_df = hed
  
)

saveRDS(hedonic_inline, "paper/output/regression/hedonic_inline.rds")
```
